Privacy-Preserving Localization Using Private Linear-Combination Aggregation

### 2020-11-27 main.tex
Disappeared brackets from Zotero auto-export re-added and auto-export disabled (outside of this repo)
All of Benjamin's final change suggestions added

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localization methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localization protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localization filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
State Estimation, Data Privacy, Sensor Fusion, Kalman Filters.
\end{IEEEkeywords}

@@

\section{Introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localization methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localization methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large applications due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localization scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighboring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localization filter using this scheme such that sensor and navigator privacy is preserved.

@@

\subsection{Notation}
Throughout this work, we will use the following notation. Lowercase underlined characters are vectors, $\vec{a}$, and random variables are bold, $\rv{a}$. Uppercase bold characters are matrices, $\mat{A}$, with $\mat{A}^{-1}$ denoting the inverse and $\mat{A}^\top$ the transpose. $|a|$ is the absolute value of $a$ and $\lVert\vec{a}\rVert$ the vector normal. Subscripts of the form $\rv{x}_{k|j}$ will denote an estimate of $\rv{x}_k$ using measurements from times up to and including $j$. $\E{\cdot}$ and $\Var{\cdot}$ are used for the expected value and variance, respectively, and encryption and decryption with key $k$ are denoted $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$. Additionally, key $k$ may be ommited when inferrable from context. $\lfloor\cdot\rceil$ rounds to the nearest integer, $a\|b$ stands for a bitwise concatenation of $a$ and $b$ and sets will be written as $\{\dots\}$ while ordered sequences as $\langle\dots\rangle$.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalizes the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

In the context of privacy-preserving localization and state estimation, some of the aforementioned schemes have found applications and some specialized encryption methods have been proposed. These have been summarized in the next section.

@@

\subsection{Applications to Localization and Estimation}
Localization methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide a computationally plausible and accurate localization filter in the presence of range-only sensors while providing uncertainties associated with estimates.

@@

Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralized or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ however, in that weights are generalized to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants.

@@

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by a different \textit{instance} variable $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarized in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localization} we will show how this protocol is sufficient to compute the required estimation information at the sensors needed to update the navigator's estimation filter.

@@

\subsection{Inherent Leakage}
Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates as is the case at the navigator, or the knowledge of one or more measurements as is the case by one or more malicious sensors, naturally leads to the leakage of some information. For this reason, we accept that average sensor information may be leaked to the navigator and that less accurate estimates of location may be leaked to sensors. We will Instead show that individual sensor information remains private to sensors and that the most precise localization estimate, requiring all measurements, remains private to the navigator.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme and the Joye-Libert private aggregation scheme.

@@

\subsection{Security Notions}
The security of a cryptographic scheme is typically defined by a security \textit{game}, which captures both the desired privacy guarantees, as well as the capabilities of attackers \cite{katzIntroductionModernCryptography2008}. The typical security notion for a homomorphic encryption scheme is Indistinguishability under Chosen Plaintext Attack (IND-CPA) \cite{chaseSecurityHomomorphicEncryption2017}. 
\begin{definition}
An encryption scheme meets IND-CPA security if an attacker who can choose plaintext messages to be encrypted at will, gains no additional information about an unknown plaintext message when they learn only its encryption.

The formal security game for IND-CPA has been given in Appendix \ref{app:ind_cpa}.
\end{definition}
Private aggregation schemes aim for the security notion of Aggregator Obliviousness (AO) \cite{shiPrivacyPreservingAggregationTimeSeries2011}. 
\begin{definition}
An encryption scheme meets AO security if no colluding subset of participants \textit{excluding} the aggregator gains additional information about the remaining aggregation values given only their encryptions, while any colluding subset \textit{including} the aggregator learns only their sum.

The formal security game for AO has been given in Appendix \ref{app:ao}.
\end{definition}

@@

\section{Private Linear-Combination Aggregation} \label{sec:lcao_definition}
As stated in Section \ref{subsec:problem_decomposition}, we decompose the localization method such that the multi-party aggregation protocol, given in Figure \ref{fig:agg_steps}, captures all communication between the navigator and sensors. To perform this protocol in a secure manner, we must first describe the security properties we aim to achieve during aggregation. Broadly speaking, we want a cryptographic scheme which allows instances of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarized by the two informal security notions:
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous instances, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear-Combination Aggregation] No colluding subset \textit{excluding} the aggregator gains additional information about the remaining sensor values to be weighted $a^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the aggregator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}

\begin{remark} \label{rem:lcao_leakage}
    The notion of a leakage function including parameters from the aggregator requires extra care to be taken when giving its definition. Since an attacker may compromise the aggregator, they have control over the choice of these parameters, and therefore over the leakage function. We note that in the leakage function above, $\sum^n_{i=n}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by the colluding subset, \textit{e.g.} $\sum^m_{j=1}a^{(t)}_{i,j}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the requirement that all sensors receive the same weights.
\end{remark}

@@

\subsection{Linear-Combination Aggregator Oblivious Encryption} \label{subsec:lcao}
We let a linear-combination aggregator oblivious encryption scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These are used by a trusted third party, an aggregator and value producing users $i,\,1\leq i \leq n$. They are defined as follows

\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the aggregator's public and private keys $pk_0$ and $sk_0$, and the user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, x)$] The aggregator and users can encrypt any value $x$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, user $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}

@@

\subsection{Our Scheme} \label{subsec:our_scheme}
Our scheme is based on the Paillier and Joye-Libert schemes introduced in Sections \ref{subsec:paillier_scheme} and \ref{subsec:joye_libert_scheme}, and similarly bases its security on the DCRA. As with Joye-Libert's private aggregation scheme, a trusted party is required for the initial distribution of user secret keys. Below, we give definitions for the four algorithms comprising the linear-combination aggregation encryption scheme.

@@

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localization methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localization protocols, thus not preserving participants’ data privacy. Existing approaches that do provide such guarantees fail to address sensor and navigator privacy in common non-linear measurement model applications and therefore forfeit broad applicability. We define a suitable notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Kalman Filter with range-sensor measurements. Our method keeps navigator location, sensors' locations, and sensors' measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localization filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
State Estimation, Data Privacy, Sensor Fusion, Extended Kalman Filters.
\end{IEEEkeywords}

@@

\section{Introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localization methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004}, and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localization methods, including the use of radar measurements, GPS clocks, and signal strength measurements, have found particularly large applications due to their robustness and accuracy. These methods use distance measurements to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location, and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localization scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighboring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows the private aggregation of weighted sums, before developing a computationally plausible localization filter in the presence of range-only sensors using this scheme, such that sensor and navigator privacy is preserved.

@@

\subsection{Notation}
Throughout this work, we will use the following notation. Lowercase underlined characters are vectors, $\vec{a}$, and random variables are bold, $\rv{a}$. Uppercase bold characters are matrices, $\mat{A}$, with $\mat{A}^{-1}$ denoting the inverse and $\mat{A}^\top$ the transpose. $|a|$ is the absolute value of $a$ and $\lVert\vec{a}\rVert$ the vector norm. Subscripts of the form $\rv{x}_{k|j}$ will denote an estimate of $\rv{x}_k$ using measurements from times up to and including $j$. $\E{\cdot}$ and $\Var{\cdot}$ are used for the expected value and variance, respectively, and encryption and decryption with key $k$ are denoted $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$. Additionally, key $k$ may be ommited when inferrable from context. $\lfloor\cdot\rceil$ rounds to the nearest integer, $a\|b$ stands for a bitwise concatenation of $a$ and $b$ and sets will be written as $\{\dots\}$ while ordered sequences as $\langle\dots\rangle$.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalizes the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their aggregation (typically sum) can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

In the context of privacy-preserving localization and state estimation, some of the aforementioned schemes have found applications and some specialized encryption methods have been proposed. These are summarized in the next section.

@@

\subsection{Applications to Localization and Estimation}
Localization methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide an accurate localization filter in the presence of range-only sensors while providing uncertainties associated with estimates.

@@

Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralized or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ, however, in that weights are generalized to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants. We define our alternative security notion in Section \ref{sec:lcao_definition}.

@@

The navigator state captures the position, and change in position ($\Delta$), and is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \Delta\rv{x} & \rv{y} & \Delta\rv{y}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependent on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by a different \textit{instance} variable, $t$. The desired bi-directional communication protocol at instance $t$ consists of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarized in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localization}, we will show how this protocol is sufficient to compute the required estimation information at the sensors needed to update the navigator's estimation filter.

@@

\subsection{Inherent Leakage}
Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates as is the case at the navigator, or the knowledge of one or more measurements as is the case by one or more malicious sensors, naturally leads to the leakage of some information. For this reason, we accept that average sensor information may be leaked to the navigator and that less accurate estimates of location may be leaked to sensors. We will instead show that individual sensor information remains private to sensors and that the most precise localization estimate, requiring all measurements, remains private to the navigator.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme, and the Joye-Libert private aggregation scheme.

@@

\subsection{Security Notions}
The security of a cryptographic scheme is typically defined by a security \textit{game}, which captures both the desired privacy guarantees, as well as the capabilities of attackers \cite{katzIntroductionModernCryptography2008}. The typical security notion for a homomorphic encryption scheme is Indistinguishability under Chosen Plaintext Attack (IND-CPA) \cite{chaseSecurityHomomorphicEncryption2017}, while for private aggregation schemes, the notion is Aggregator Obliviousness (AO) \cite{shiPrivacyPreservingAggregationTimeSeries2011}. 
\begin{definition}
An encryption scheme meets IND-CPA security if an attacker who can choose plaintext messages to be encrypted at will, gains no additional information about an unknown plaintext message when they learn only its encryption. The formal security game and definition for IND-CPA has been given in Appendix \ref{app:ind_cpa}.
\end{definition}
\begin{definition}
An encryption scheme meets AO security if no colluding subset of participants \textit{excluding} the aggregator gains additional information about the remaining aggregation values given only their encryptions, while any colluding subset \textit{including} the aggregator learns only their sum. The formal security game and definition for AO has been given in Appendix \ref{app:ao}.
\end{definition}

The Paillier encryption scheme and Joye-Libert private aggregation scheme meet the above security notions, respectively, and have been given below.

@@

\section{Private Linear-Combination Aggregation} \label{sec:lcao_definition}
As stated in Section \ref{subsec:problem_decomposition}, we decompose the localization method such that the multi-party aggregation protocol, given in Figure \ref{fig:agg_steps}, captures all communication between the navigator and sensors. To perform this protocol in a secure manner, we must first describe the security properties we aim to achieve during aggregation. Broadly speaking, we want a cryptographic scheme which allows instances of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarized by the two informal security notions:
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous instances, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear-Combination Aggregation] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}

\begin{remark} \label{rem:lcao_leakage}
    The notion of an aggregator leakage function including parameters from the aggregator requires extra care to be taken when giving its definition. Since an attacker may compromise the navigator (in our case the aggregator), they have control over the choice of these parameters, and therefore over the leakage function. We note that in the leakage function above, $\sum^n_{i=n}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by the colluding subset, \textit{e.g.}, $\sum^m_{j=1}a^{(t)}_{i,j}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the requirement that all sensors receive the same weights.
\end{remark}

@@

\subsection{Linear-Combination Aggregator Oblivious Encryption} \label{subsec:lcao}
We let a linear-combination aggregator oblivious encryption scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These are used by a trusted third party, an aggregator, and value producing users $i,\,1\leq i \leq n$. They are defined as follows

\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the aggregator's public and private keys $pk_0$ and $sk_0$, and the user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, x)$] The aggregator and users can encrypt any value~$x$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, user $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}

@@

\subsection{Our Scheme Meeting LCAO Security} \label{subsec:our_scheme}
Our scheme is based on the Paillier and Joye-Libert schemes introduced in Sections \ref{subsec:paillier_scheme} and \ref{subsec:joye_libert_scheme}, and similarly bases its security on the DCRA. As with Joye-Libert's private aggregation scheme, a trusted party is required for the initial distribution of user secret keys. Below, we give definitions for the four algorithms comprising the linear-combination aggregation encryption scheme.

@@

\begin{definition}
    An encryption scheme meets IND-CPA security if no adversary, running in probabilistic-time with respect to security parameter $\kappa$, has more than a negligible advantage in winning the above security game. Probabilities are taken over randomness introduced by $\mathcal{A}$, and in $\mathsf{Setup}$ and $\mathsf{Enc}$.
\end{definition}
